### 2024-04-19 docs/SOSP 2024/abstract.tex
Update abstract. Add page numbers

  This paper is motivated by the growing demand for the increased performance
  of machine learning applications on different hardware platforms
  including CPUs and GPUs. We focus on accelerating the inference of decision tree based models, 
  which are the most popular models for tabular data. Existing solutions do not achieve 
  the highest possible performance because they do not explore different optimization
  configurations. And since these systems are hand-written, they are not portable either.

  We address these problems by designing \Treebeard{}, a \emph{schedule-guided},
  \emph{retargetable} compiler infrastructure for decision tree based models.  
  \Treebeard{} has two core components. The first is a scheduling language 
  that encapsulates the large optimization space for decision tree inference, 
  and techniques to efficiently explore this space. \TODO{Change large optimization space}.
  The second is an optimizing retargetable multi-level compiler that 
  can generate code for any specified schedule. \Treebeard{}'s retargetability 
  is based not only on being able to generate code for different target architectures (CPU vs. GPU),
  but also on its ability to use different data layouts, caching strategies,
  parallel reduction schemes etc. To accomplish this level of configurability, we re-architect 
  and significantly extend the open-source \TreebeardOLD{} CPU compiler
  to support (i) schedule-guided compilation, (ii) retargetable GPU code
  generation, and (iii) GPU-specific optimizations. 
  %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.

  We demonstrate that \Treebeard{} can generate high-performance inference 
  code, for several hundred decision tree models across different batch sizes
  and target architectures. Our scheduling heuristic is able to quickly find near-optimal schedules 
  \TODO{[how do we argue that the schedule is near-optimal? Do we have some 
  exptl. evidence that we can show in the results section?]}
  schedule while searching over a small number ($\sim$50) of schedules.
  In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only 
  target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
  On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.
  For models where \TreebeardOLD{} was only able to achieve diminishing returns with an 
  increasing number of threads, \Treebeard{} is able to scale linearly with the number of threads.
  \TODO{(numbers for CPU performance?)}
\end{abstract}

>>>>

  The rapid proliferation of machine learning together with the accelerating evolution of 
  the hardware ecosystem has led to a surge in the demand for model inference on a variety
  of hardware. This paper is motivated by the problems encountered when targeting 
  inference of decision tree based models, the most popular models on tabular data,
  to run at peak performance on diverse CPU and GPU targets. Existing solutions are neither
  portable nor achieve the best possible performance for the specific hardware they 
  target because they do not explore different optimization configurations.

  We present \Treebeard{}, a \emph{schedule-guided},
  \emph{retargetable} compiler for decision tree based models
  that searches over several optimization choices and automatically generates 
  high-performance inference routines for CPUs and GPUs.  
  \Treebeard{} has two core components. The first is a scheduling language 
  that encapsulates the optimization space for decision tree inference, 
  and techniques to efficiently explore this space. 
  The second is an optimizing retargetable compiler that 
  can generate code for any specified schedule. \Treebeard{}'s ability to use different data layouts, 
  loop structures and caching strategies enables it to achieve portable performance.   
  % To accomplish this level of configurability, we re-architect 
  % and significantly extend the open-source \TreebeardOLD{} CPU compiler
  % to support (i) schedule-guided compilation, (ii) retargetable GPU code
  % generation, and (iii) GPU-specific optimizations. 
  %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.

  We evaluate \Treebeard{} on several hundred decision tree models across different batch sizes
  and target architectures. We find that our schedule exploration strategy is able to quickly
  find near-optimal schedules.
  In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
  about 2-5$\times$ faster on average than RAPIDS FIL and Tahoe over several batch sizes.
  While these systems only target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
  On CPUs, \Treebeard{} is able to outperform \TreebeardOLD{} by up to $5\times$ by utilizing 
  additional sources of parallelism at small batch sizes.
\end{abstract}
